---
title: "Assignment 2"
Author: Shawn Ye
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(leaflet)
library(tidyverse)
```

## Data Wrangling
```{r}
ind <- data.table::fread("/Users/SF/Desktop/Study/PM566/chs_individual.txt")
reg <- data.table::fread("/Users/SF/Desktop/Study/PM566/chs_regional.txt")
merg <- merge(x = ind, y = reg, by = "townname", all.x = TRUE, all.y = FALSE)
merg[,mbmi:=mean(bmi,na.rm=TRUE),by=.(male, hispanic)]
merg[,mfev:=mean(fev,na.rm=TRUE),by=.(male, hispanic)]
merg[,bmi:=ifelse(is.na(bmi),mbmi,bmi)]
merg[,fev:=ifelse(is.na(fev),mfev,fev)]
merg[,mgasstove := round(mean(gasstove,na.rm=TRUE)),by=.(male,hispanic)]
merg[,msmoke := round(mean(smoke,na.rm=TRUE)),by=.(male,hispanic)]
merg[,gasstove:=ifelse(is.na(gasstove),mgasstove,gasstove)]
merg[,smoke:=ifelse(is.na(smoke),msmoke,smoke)]
```

##1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.
```{r}
nrow(merg) == nrow(ind)
```

##2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r}
merg[bmi<14, obesity_level := "underweight"]
merg[bmi>=14 & bmi<22, obesity_level := "normal"]
merg[bmi>=22 & bmi<24, obesity_level := "overweight"]
merg[bmi>=24, obesity_level := "obese"]
merg[,.(
  bmi_min = min(bmi, na.rm=T),
  bmi_max = max(bmi, na.rm=T),
  n= length(bmi)
),by =  obesity_level] [order( obesity_level)]
```

##3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
```{r}
merg[smoke==0 & gasstove ==0, smoke_gas_exposure := "No exposure"]
merg[smoke==1 & gasstove ==0, smoke_gas_exposure := "Smoke exposure"]
merg[smoke==0 & gasstove ==1, smoke_gas_exposure := "Gasstove exposure"]
merg[smoke==1 & gasstove ==1, smoke_gas_exposure := "Both exposure"]
merg[,.(
   n= length(smoke)
), by = smoke_gas_exposure][order( smoke_gas_exposure)]
```

##4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r}
merg[,.(
   fev_mean = mean(fev, na.rm=T),
   fev_sd = sd(fev, na.rm=T), 
   asthma_prop = mean(asthma, na.rm=T),
   asthma_sd = sd(asthma,na.rm=T)), by = townname][order(townname)]

merg[,.(
   fev_mean = mean(fev, na.rm=T),
   fev_sd = sd(fev, na.rm=T),
   asthma_prop = mean(asthma, na.rm=T),
   asthma_sd = sd(asthma,na.rm=T)), by = male][order(male)]

merg[,.(
   fev_mean = mean(fev, na.rm=T),
   fev_sd = sd(fev, na.rm=T),
   asthma_prop = mean(asthma, na.rm=T),
   asthma_sd = sd(asthma,na.rm=T)), by = obesity_level][order(obesity_level)]

merg[,.(
   fev_mean = mean(fev, na.rm=T),
   fev_sd = sd(fev, na.rm=T),
   asthma_prop = mean(asthma, na.rm=T),
   asthma_sd = sd(asthma,na.rm=T)), by = smoke_gas_exposure][order(smoke_gas_exposure)]
```


## Looking at the Data (EDA)
##1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}
merg %>%
  filter(!(townname %in% NA)) %>%
  ggplot(mapping = aes(x= bmi, y=fev, color=townname))+
  geom_point()+facet_wrap(~townname)+geom_smooth(method = "lm")+
  labs(title="Regression lines of BMI vs FEV by townname", x="BMI",y="FEV")
```

It can be seen from the graphs that there is a positive association between BMI and FEV through all the towns.

##2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
merg %>%
  filter(!(obesity_level %in% NA)) %>% 
  ggplot()+
  geom_histogram(mapping = aes(x=fev,fill=obesity_level))+
  labs(title = "FEV Values by BMI Category",x="FEV")+
  scale_fill_brewer(palette="Accent")

merg %>%
  filter(!(smoke_gas_exposure %in% NA)) %>% 
  ggplot()+
  geom_histogram(mapping = aes(x=fev,fill=smoke_gas_exposure))+
  labs(title = "FEV Values by smoke/gas exposure",x="FEV")+
  scale_fill_brewer(palette="Paired")
```

It can be seen that overweight and obese children tend to have higher FEV, while underweight children tend to have lower FEV, compared with children with normal weight. Dissimilarly, FEV does not show strong association with smoke and gas exposure status based on the histograms.

##3. Barchart of BMI by smoke/gas exposure.
```{r}
merg %>% 
  filter(!(smoke_gas_exposure %in% NA)) %>% 
  ggplot()+
  geom_bar(mapping = aes(x = obesity_level, fill = smoke_gas_exposure))+
  labs(title = "BMI Category by Smoke and Gas Exposure")+
  scale_fill_brewer(palette = "Pastel2")
```

The barchart shows that gasstove exposure accounts for the most in children with all kinds of obesity level.

##4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
merg %>%
  filter(!(obesity_level %in% NA)) %>%
  ggplot(mapping=aes(x=obesity_level, y=fev))+
  stat_summary(fun.min=min, fun.max=max, fun=mean)
  labs(title = "Average FEV by BMI Category")
  
merg %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
  ggplot(mapping=aes(x=smoke_gas_exposure, y=fev))+
  stat_summary(fun.min=min, fun.max=max, fun=median)
  labs(title = "Average FEV by smoke/gas exposure category")
```

Again, we noticed that obese and overweight children have higher mean FEV, and underweight children have lower FEV, compared with children with normal weight. Smoke and gas exposure do not seem to be associated with FEV, which is consistent with the previous conclusion.

##5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
pm25_pal = colorNumeric(c('darkgreen','goldenrod','brown'), domain=merg$pm25_mass)
leaflet(merg)%>%
  addProviderTiles("OpenStreetMap")%>%
  addCircles(lat=~lat,lng=~lon,color=~pm25_pal(pm25_mass),opacity=1,fillOpacity=1, radius=500)%>%
addLegend("bottomleft", pal=pm25_pal,values =merg$pm25_mass, title="concentration of PM2.5 Mass in each of the CHS communities", opacity=1)
```

##6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}
merg %>% 
  filter(!(pm25_mass %in% NA)) %>%
  ggplot(mapping = aes(x = pm25_mass,y = fev))+
  geom_point()+
  geom_smooth()
```

It can be seen from the graph that there is a weak negative association between FEV and PM2.5 mass.
